Evolution of Sex-Specific Recombination Rate

### 2020-05-14 doc/MS_v1.Rmd
extensive debugging and organizing the datascripts

---
title: 'Evolution of Sex-Specific Recombination Rate'
date: '`r Sys.Date()`'
author: "April L. Peterson, Bret Payseur"
output:
  word_document:
    toc: yes
    toc_depth: '5'
  pdf_document:
    toc: yes
    toc_depth: '5'
  html_document:
    toc: yes
    toc_depth: '5'
    toc_float: yes
always_allow_html: yes
bibliography: refs_Main_Report.bib
---

@@

load(file="~./MLH1repo/data/Results_setup_4.24.20.RData")

load(file="~./MLH1repo/data/Results_setup_4.21.20.RData")

load(file="~./MLH1repo/data/CleanBivData_6.3.20.RData")
load(file="~./MLH1repo/data/total_SC_30.12.19.RData")

#skeletonize
SC.skel = read.csv("~./MLH1repo/data/SCskel_output_feb20.csv", header = TRUE, strip.white = TRUE)

@@

To account for confounding effects of sex chromosomes from pooled samples of bivalents, we considered a **reduced data** set including only bivalents with SC lengths below the 2nd quartile in cells with at least 17 of 20 single bivalent measures. The 'short bivalent' data set included the four or five shortest bivalents, and excluded the X bivalent in oocytes. A total of `r sum(Q1.real.short_mouse_table$n_obs)` 'short' bivalents were isolated from `r sum(  (Q1.real.short_mouse_table  %>% filter(sex=="female") )$ncells)` oocytes and `r sum(  (Q1.real.short_mouse_table  %>% filter(sex=="male") )$ncells)` spermatocytes. Although this smaller data set has decreased power, it offers a more comparable set of single bivalents to compare between the sexes.  A 'long bivalent' data set was calculated as above the 4th quartille in SC lengths per cell.

@@

#MLH1_data <- with(MLH1_data, MLH1_data[order(category),])
AP_mouse_table.HQ$var <- as.numeric(AP_mouse_table.HQ$var)

AP_mouse_table.HQ_target <- AP_mouse_table.HQ[!(is.na(AP_mouse_table.HQ$sex) | AP_mouse_table.HQ$sex==""), ]

@@

```{r MLH1.MIXED.MODEL.setup, echo=FALSE, warning=FALSE, eval=TRUE}
#DF.HetC.MixedModel.HQ
#this blck is duplicated in setupData.rm

#https://rcompanion.org/handbook/G_03.html
#testing random effects within lme

DF.HetC.MixedModel <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$species == "M.musculus",]

#AP_mouse_table excluide Qual5 cells
DF.HetC.MixedModel.HQ <- AP_mouse_table.MLH1_HQ
DF.HetC.MixedModel.HQ <- add_species(DF.HetC.MixedModel.HQ)

@@

@@

@@

female_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
                            subsp.y,subsp.x,
                    category.y,category.x, strain.x,
                    strain.y, mouse.1  )  )

@@

#+theme(axis.text.x = element_text(angle = 45, hjust = 1))

@@

```{r MLH1.sum.stats.Bret.edits, echo=FALSE}

AP_mouse_table.HQ <- add_species(AP_mouse_table.HQ)
#total mice
#length( unique(AP_mouse_table.HQ %>% filter(species == "M.musculus") )$mouse )

#average cell number for spermatocytes and oocytes
#mean(female_mouse.avs_for.analysis$Ncells, na.rm = T)
#mean( (male_mouse.av %>% filter(species == "M.musculus") )$Ncells, na.rm = T)
```

We used counts of MLH1 foci per cell to estimate genome-wide recombination rates in `r length(unique(  (main.table  %>% filter(species == "M.musculus") )$strain) )` wild-derived inbred strains sampled from three subspecies of house mice: *M. m. domesticus*, *M. m. musculus* and *M. m. molossinus*. Mean MLH1 foci counts for `r length( unique(AP_mouse_table.HQ_target)$mouse )` mice were quantified from an average of `r mean( (AP_mouse_table.HQ_target %>% filter(sex == "male") )$Ncells, na.rm = T)` spermatocytes per male (for a total of `r formatC(sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "male"))$Ncells),  format="d", big.mark=",")` spermatocytes) and `r mean(female_mouse.avs_for.analysis$Ncells, na.rm = T)` oocytes per female (for a total of `r formatC(sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "female"))$Ncells), format="d", big.mark=",")` oocytes).

@@

```{r DMC1.setup, echo=FALSE}
#load DMC1 csv
#DMC1.DF <- read.csv("~./DMC1/newJuviQuant.csv", header = TRUE)
DMC1.DF <- read.csv("~./MLH1repo/data/DMC1newJuviQuant.csv", header = TRUE)

DMC1.DF$fileName <- DMC1.DF$Original.Name
DMC1.DF <- add_mouse(DMC1.DF)
DMC1.DF <- add_strain(DMC1.DF)
DMC1.DF <- add_category(DMC1.DF)
DMC1.DF <- add_subsp(DMC1.DF)
DMC1.DF <- DMC1.DF[DMC1.DF$stage != "P",]

#make DMC1 table
DMC1.mean.obs.L <- ddply(.data=DMC1.DF[DMC1.DF$stage == "Early",],
                 .(category),
                 summarize,

                 ncells = length(fileName),
                 mean.count = mean( total.count)
)
DMC1.mean.obs.L$stage <- "early Z"

DMC1.mean.obs.Z <- ddply(.data=DMC1.DF[DMC1.DF$stage == "Late",],
                 .(category),
                 summarize, 
                 ncells = length(fileName),
                 mean.count = mean( total.count)
)
DMC1.mean.obs.Z$stage <- "late Z"

DMC1.mean.table <- rbind(DMC1.mean.obs.L,DMC1.mean.obs.Z)

DMC1.DF$rec_group <- ifelse(grepl("WSB", DMC1.DF$strain), "Low",
                          ifelse(grepl("G",DMC1.DF$strain), "Low",

              ifelse(grepl("KAZ", DMC1.DF$strain), "Low",
            ifelse(grepl("PWD", DMC1.DF$strain), "High",
               ifelse(grepl("MSM", DMC1.DF$strain), "High","")))))
```

@@

#make category averages from AP_mouse_table.HQ
mean.MLH1.obs <- ddply(.data=AP_mouse_table.HQ,
                 .(category),
                 summarize, 
                 mean.MLH1 = mean(mean_co),
                 var.MLH1 = var(mean_co)
)

@@

chrm.class.barplot
kk <- addSmallLegend(chrm.class.barplot)

@@

#strain av tables
#chrm.prop.male.high1
#chrm.prop.male.low
```

@@

```{r Q1.Short.Long_biv.setup, echo=FALSE}

#make the Q1 strain list
Q1.real.short_mouse_table <- REAL.short_mouse_table %>% filter( (strain != "CZECH") & (strain != "MOLF") )

Q1.real.long_mouse_table <- REAL.long_mouse_table %>% filter( (strain != "CZECH") & (strain != "MOLF") )

```{r Q1.show.SC.length, echo=FALSE, warning=FALSE}

@@

```{r Q1.SC.show.short.biv, echo=FALSE}

Q1.SC.short.mouse_scatter_boxplot <- ggplot( Q1.real.short_mouse_table, aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(20,115) )+ggtitle("")+  theme(legend.position="none", axis.title.x = element_blank(), axis.text.x = element_text(size=9), axis.title.y = element_text(size = 10), axis.text.y = element_text(size=9))+labs(y = "Mouse Mean Short SC Length")

@@

#run t.tests for the mouse av. for reduced single bivalent datasets
#WSB
Q1_mus.av_shortSC_t.test.WSB <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC,
                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC)

#G
Q1_mus.av_shortSC_t.test.G <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC,
                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC)

#LEW
Q1_mus.av_shortSC_t.test.LEW <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean_SC,
                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean_SC)

#PWD
Q1_mus.av_shortSC_t.test.PWD <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean_SC,
                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean_SC)

#KAZ
Q1_mus.av_shortSC_t.test.KAZ <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean_SC,
                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean_SC)

#SKIVE
Q1_mus.av_shortSC_t.test.SKIVE <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean_SC,
                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean_SC)

#MSM
#Q1_mus.av_shortSC_t.test.MSM <-t.test( (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "MSM") )$mean_SC, 
#          (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == #"MSM"))$mean_SC )
#1 male, 3 females measures
#try to merge the manual - whole cell observations (at least 2 males)

```{r Q1SC_short.biv_dif, echo=FALSE}
#calq the different in short.biv means across sexes

#WSB
Q1SC.mus.av.short.biv_DIFF_WSB <- mean( (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC) / mean(  (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC )

#LEW
Q1SC.mus.av.short.biv_DIFF_LEW <- mean( (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean_SC) / mean(  (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean_SC )

#G
Q1SC.mus.av.short.biv_DIFF_G <- mean( (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC) / mean(  (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC )

#PWD
Q1SC.mus.av.short.biv_DIFF_PWD <- mean( (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean_SC) / mean(  (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean_SC )

#KAZ
Q1SC.mus.av.short.biv_DIFF_KAZ <- mean( (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean_SC) / mean(  (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean_SC )

#SKIVE
Q1SC.mus.av.short.biv_DIFF_SKIVE <- mean( (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean_SC) / mean(  (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean_SC )

#MSM
Q1SC.mus.av.short.biv_DIFF_MSM <- mean( (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "MSM") )$mean_SC) / mean(  (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "MSM") )$mean_SC )

#this bivData ratio is the same as mouse average
Q1SC.BivData.short.biv_DIFF_MSM <- mean( (real.short.bivData %>% filter(sex == "female") %>% filter(strain == "MSM") )$chromosomeLength) / mean(  (real.short.bivData %>% filter(sex == "male") %>% filter(strain == "MSM") )$chromosomeLength )

@@

#change this plot to look like the short bivalent plot -- boxplots + mouse points
Tot.SC_mouse.av_ALL <- ggplot(data= Q1.totSC.mouse.av.table, aes(y = mean.skel, x=sex, color=strain))+
  scale_color_manual(values=colors_of_strains)+

  facet_wrap(~subsp, scales="free", ncol = 3)+ 
  geom_jitter()+ geom_boxplot(alpha=.2)+ 
  #geom_hline(data =totSC.strain.av.table, aes(yintercept=mean.skel) )+

  theme(axis.title.x=element_blank(), axis.ticks.x=element_blank(),axis.text.x = element_text(size=9), legend.position="none", axis.text.y = element_text(size=9), axis.title.y = element_text(size = 10)  )+  
  labs(y = "Mouse Mean Total SC Length")+
  ylim(c(1000,2600 ))+ggtitle("")

@@

two.Q1.SC.plots <- {Q1.SC.short.mouse_scatter_boxplot + {Tot.SC_mouse.av_ALL} + plot_layout(ncol = 2, widths = c(1,1), heights = c(2,2)) }

@@

@@

```{r Q1.M1_short_biv_lmer, echo=FALSE, warning=FALSE, message=FALSE}
lmer.Q1_SCLength_short_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE)

lmer.Q1_SCLength_short_M1_R.sex <- lmer(mean_SC ~ subsp + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE)

lmer.Q1_SCLength_short_M1_R.subsp <- lmer(mean_SC ~ sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE)

@@

We use the reduce data set of short bivalents to remove the effects of the paired X-X bivalent in females. The mouse average of short bivalents are significantly longer in females than males in all strains (t-test; p < 0.05) except _musculus^SKIVE^_, which my be attributed to low sample sizes (p = `r Q1_mus.av_shortSC_t.test.SKIVE$p.value`). The ratios of the mouse mean SC across sexes range across strains from `r Q1SC.mus.av.short.biv_DIFF_MSM` ( _musculus^MSM^_ ) to `r Q1SC.mus.av.short.biv_DIFF_WSB` ( _domesticus^WSB^_ ).

@@

Q1.real.short_mouse_table$subsp <- factor(Q1.real.short_mouse_table$subsp, ordered = FALSE)
Q1.real.short_mouse_table$sex <- factor(Q1.real.short_mouse_table$sex, ordered = FALSE)
Q1.real.short_mouse_table$strain <- factor(Q1.real.short_mouse_table$strain, ordered = FALSE)

Q1.glm_short.SC_M2 <- glm(mean_SC ~ subsp * strain * sex, data=Q1.real.short_mouse_table)

@@

#round(summ.Q1.glm_short.SC_M2$coefficients[,c(1,4)],5)

@@

Q1.glm_short.SC_M3 <- glm(mean_SC ~ strain * sex, data=Q1.real.short_mouse_table)

@@

#make comparisons for the high quality cells

#get Q12 cells from MLH1.merge.Skel
#calq the mouse averages

@@

totSC_Q12_mouse.av.table <- ddply(.data=MLH1.merge.Skel %>% filter(quality < 3), 
                 .(mouse),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                # mean.MLH1 = mean( nMLH1.foci, na.rm = TRUE),
                # var.MLH1 = var( nMLH1.foci, na.rm = TRUE),

@@

#mouse av
Q1.SC.long.mouse_scatter_boxplot <- ggplot( Q1.real.long_mouse_table, aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(75,150) )+ggtitle("Mouse average SC Length \n long bivalent dataset")

```{r Q1_SC.biv.long_ttest, echo=FALSE, include=FALSE}

@@

```{r Q1.F1pos_long.short_setup, echo=FALSE}

@@

```{r Comp_Ratios, echo=FALSE}

>>>>

---
title: 'Evolution of Sex-Specific Recombination Rate'
date: '`r Sys.Date()`'
author: "April L. Peterson, Bret Payseur"
output:
  html_document:
    toc: yes
    toc_depth: '5'
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '5'
  word_document:
    toc: yes
    toc_depth: '5'
always_allow_html: yes
bibliography: refs_Main_Report.bib
---

@@

load(file="~./MLH1repo/data/Results_setup_5.14.20.RData") #Results_setup_5.14.20.RData

#load(file="~./MLH1repo/data/Results_setup_4.21.20.RData")

#load(file="~./MLH1repo/data/CleanBivData_6.3.20.RData")

#
#load(file="~./MLH1repo/data/total_SC_30.12.19.RData")

#skeletonize - this is the total sc file
#SC.skel = read.csv("~./MLH1repo/data/SCskel_output_feb20.csv", header = TRUE, strip.white = TRUE)

@@

To account for confounding effects of sex chromosomes from pooled samples of bivalents, we considered a **reduced data** set including only bivalents with SC lengths below the 2nd quartile in cells with at least 17 of 20 single bivalent measures. The 'short bivalent' data set included the four or five shortest bivalents, and excluded the X bivalent in oocytes. A total of `r sum(Long.biv_mouse_table$n_obs)`   'short' bivalents were isolated from `r sum(  (Long.biv_mouse_table  %>% filter(sex=="female") )$ncells)` oocytes and `r sum(  (Long.biv_mouse_table  %>% filter(sex=="male") )$ncells)` spermatocytes. Although this smaller data set has decreased power, it offers a more comparable set of single bivalents to compare between the sexes.  A 'long bivalent' data set was calculated as above the 4th quartille in SC lengths per cell.

@@

#MLH1_data <- with(MLH1_data, MLH1_data[order(category),])
Mouse.Table_HQ$var <- as.numeric(Mouse.Table_HQ$var)

AP_mouse_table.HQ_target <- Mouse.Table_HQ[!(is.na(Mouse.Table_HQ$sex) | Mouse.Table_HQ$sex==""), ]

@@

```{r MLH1.MIXED.MODEL.setup, echo=FALSE, warning=FALSE, eval=TRUE}

#DF.HetC.MixedModel.HQ
#this blck is duplicated in setupData.rm
#https://rcompanion.org/handbook/G_03.html
#testing random effects within lme

DF.HetC.MixedModel <- Mouse.Table_HQ[Mouse.Table_HQ$species == "M.musculus",]

#AP_mouse_table excluide Qual5 cells
DF.HetC.MixedModel.HQ <- Mouse.Table_HQ
DF.HetC.MixedModel.HQ <- add_species(DF.HetC.MixedModel.HQ)

@@

#
strain.table.HQ_Fig.og <- add_subsp(strain.table.HQ_Fig.og)
strain.table.HQ_Fig.og <- add_species(strain.table.HQ_Fig.og)

@@

sperman.corr.FM <- cor( ( strain.table.HQ_Fig %>% filter(sex == "female") %>% filter(species == "M.musculus" ) )$mean_co, ( strain.table.HQ_Fig %>% filter(sex == "male") %>% filter(species == "M.musculus" ) )$mean_co, 
                        method = c("spearman") )

sperman.corr.FM.pval <- cor.test( ( strain.table.HQ_Fig %>% filter(sex == "female") %>% filter(species == "M.musculus" ) )$mean_co, ( strain.table.HQ_Fig %>% filter(sex == "male") %>% filter(species == "M.musculus" ) )$mean_co, 
                        method = c("spearman") )

#write.xlsx2(strain.table.HQ_Fig, "Main_MLH1_table2.xlsx", sheetName = "Sheet1", 
#            col.names = TRUE, row.names = FALSE, append = FALSE)

#qk.plot <- plot( ( strain.table.HQ_Fig %>% filter(sex == "female") )$mean_co,
#      ( strain.table.HQ_Fig %>% filter(sex == "male") )$mean_co)

@@

female_mouse.av <- Mouse.Table_HQ %>% filter(sex == "female")

@@

#+theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

@@

We used counts of MLH1 foci per cell to estimate genome-wide recombination rates in `r length(unique(  (main.table  %>% filter(species == "M.musculus") )$strain) )` wild-derived inbred strains sampled from three subspecies of house mice: *M. m. domesticus*, *M. m. musculus* and *M. m. molossinus*. Mean MLH1 foci counts for `r length( unique(Mouse.Table_HQ)$mouse )` mice were quantified from an average of `r mean( (Mouse.Table_HQ %>% filter(sex == "male") )$Ncells, na.rm = T)` spermatocytes per male (for a total of `r formatC(sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "male"))$Ncells),  format="d", big.mark=",")` spermatocytes) and `r mean(female_mouse.avs_for.analysis$Ncells, na.rm = T)` oocytes per female (for a total of `r formatC(sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "female"))$Ncells), format="d", big.mark=",")` oocytes).

@@

@@

#make category averages from AP_mouse_table.HQ
mean.MLH1.obs <- ddply(.data=Mouse.Table_HQ,
                 .(category),
                 summarize, 
                 mean.MLH1 = mean(mean_co),
                 var.MLH1 = var(mean_co)
)

@@

kk <- addSmallLegend(chrm.class.barplot)

@@

@@

```{r Q1.Short.Long_biv.setup, echo=FALSE, eval=FALSE}
#moved this to setupdata

#REAL.long_mouse_table
#REAL.short_mouse_table
#Long.biv_mouse_table

#make the Q1 strain list
Q1.real.short_mouse_table <- Long.biv_mouse_table %>% filter( (strain != "CZECH") & (strain != "MOLF") )

Q1.real.long_mouse_table <- Long.biv_mouse_table %>% filter( (strain != "CZECH") & (strain != "MOLF") )

```{r Q1.show.SC.length, echo=FALSE, warning=FALSE, eval=FALSE}
Curated_BivData99 <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
Curated_BivData99 <- Curated_BivData99[(!is.na(Curated_BivData99$sex)),]

@@

```{r Q1.SC.show.short.biv, echo=FALSE, eval=TRUE}

#Short.biv_mouse_table

Q1.SC.short.mouse_scatter_boxplot <- ggplot( Short.biv_mouse_table, aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(20,115) )+ggtitle("")+  
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"),  legend.position="none", axis.title.x = element_blank(), axis.text.x = element_text(size=7), axis.title.y = element_text(size = 10), axis.text.y = element_text(size=9) )+

  labs(y = "Short SC Length")

@@

#run t.tests for the mouse av. for reduced single bivalent datasets

#WSB
Q1_mus.av_shortSC_t.test.WSB <- t.test((Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC,
                           (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC)

#G
Q1_mus.av_shortSC_t.test.G <- t.test((Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC,
                           (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC)

#LEW
Q1_mus.av_shortSC_t.test.LEW <- t.test((Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean_SC,
                           (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean_SC)

#PWD
Q1_mus.av_shortSC_t.test.PWD <- t.test((Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean_SC,
                           (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean_SC)

#KAZ
Q1_mus.av_shortSC_t.test.KAZ <- t.test((Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean_SC,
                           (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean_SC)

#SKIVE - this value used in the text
#only one SKVIE female, this is breaking the test
#Q1_mus.av_shortSC_t.test.SKIVE <- t.test(

#      (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean_SC,
#      (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean_SC
#      )

```{r Q1SC_short.biv_dif, echo=FALSE, eval=TRUE}

#calq the different in short.biv means across sexes
#Long.biv_mouse_table

#WSB
Q1SC.mus.av.short.biv_DIFF_WSB <- mean( (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC) / mean(  (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC )

#LEW
Q1SC.mus.av.short.biv_DIFF_LEW <- mean( (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean_SC) / mean(  (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean_SC )

#G
#Q1SC.mus.av.short.biv_DIFF_G <- mean( (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC) / mean(  (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC )

#PWD
Q1SC.mus.av.short.biv_DIFF_PWD <- mean( (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean_SC) / mean(  (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean_SC )

#KAZ
Q1SC.mus.av.short.biv_DIFF_KAZ <- mean( (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean_SC) / mean(  (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean_SC )

#SKIVE
Q1SC.mus.av.short.biv_DIFF_SKIVE <- mean( (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean_SC) / mean(  (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean_SC )

#MSM
Q1SC.mus.av.short.biv_DIFF_MSM <- mean( (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "MSM") )$mean_SC) / mean(  (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "MSM") )$mean_SC )

#this bivData ratio is the same as mouse average
Q1SC.BivData.short.biv_DIFF_MSM <- mean( (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "MSM") )$chromosomeLength) / mean(  (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "MSM") )$chromosomeLength )

@@

Q1.totSC.mouse.av.table$subsp2 <- 
   ifelse(grepl("Dom", Q1.totSC.mouse.av.table$subsp), "M. m. domesticus", 
        ifelse(grepl("Musc",Q1.totSC.mouse.av.table$subsp), "M. m. musculus",
            ifelse(grepl("Mol",Q1.totSC.mouse.av.table$subsp), "M. m. molossinus",   
               "other")))

#change this plot to look like the short bivalent plot -- boxplots + mouse points

Tot.SC_mouse.av_ALL <- ggplot(data= Q1.totSC.mouse.av.table, aes(y = mean.skel, x=sex, color=strain))+
  scale_color_manual(values=colors_of_strains)+
    facet_wrap(~subsp, scales="free", ncol = 3) +

  geom_jitter()+ geom_boxplot(alpha=.2)+ 
  #geom_hline(data =totSC.strain.av.table, aes(yintercept=mean.skel) )+

  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"),
    axis.title.x=element_blank(), axis.ticks.x=element_blank(),
axis.text.x = element_text(size=7), 
legend.position="none", axis.text.y = element_text(size=9), axis.title.y = element_text(size = 10) ) +

  labs(y = "Total SC Length")+ylim(c(1000,2600 ))+ggtitle("")

#italizing the facet label semms too hard -- do it in the final AI version
#levels(powerDF$alph) <- c("italic('p=')*0.001", "italic('p=')*0.01", "italic('p=')*0.05")

```{r Fig3.load.icons, echo=FALSE}
# load in the icons

all.chrm.yellow <- readPNG('~/MLH1repo/doc/figureFiles/all_yellow_binary.png')
single.chrm.yellow <- readPNG('~/MLH1repo/doc/figureFiles/single_chrm_yellow.png')

#EMPTY GGPLOT
all.chrm.icon <- ggplot(empty.DF, aes(x = col1, y = col2)) +
  background_image(all.chrm.yellow) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(),
    axis.title.x=element_blank(),                       axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.title.y=element_blank(),
                                  axis.text.y=element_blank(),axis.ticks.y=element_blank() )

single.chrm.icon <- ggplot(empty.DF, aes(x = col1, y = col2)) +
  background_image(single.chrm.yellow) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(),
    axis.title.x=element_blank(),
   axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.title.y=element_blank(),
                                  axis.text.y=element_blank(),axis.ticks.y=element_blank() )

```{r Fig3.example.double.triangle, echo=FALSE}
#load an example two triangle plot

single.chrm.yellow <- readPNG('~/MLH1repo/doc/figureFiles/single_chrm_yellow.png')

@@

#use cowplot
library(cowplot)

Q2.SC_dif.Figure3  <- ggdraw() +

  draw_plot(Q1.SC.short.mouse_scatter_boxplot, x=0, y=.38, width=.5, height= .6)+
  draw_plot(Tot.SC_mouse.av_ALL, x=.475, y=.38,  width = .5, height = .6)+

  draw_plot(single.chrm.icon, x=.44, y=.79,  width = .1, height = .1)+   
  draw_plot(all.chrm.icon, x=.88, y=.79,  width = .1, height = .1)+

  #drawing triangles in the corner
  #top right-triangle
  draw_line(
    x = c(.1, .2,  .1, .1 ), #keep the x values consistant
    y = c( .25, .25,  .15, .25),
    color = "blue", size=.75
  ) +

  #bottom right-triangle
  draw_line(
    x = c( .1, .2, .2, .1 ),
    y = c( .1,  .2, .1, .1),
    color = "black", size=.75
  ) +

  draw_plot_label(label = c("A", "B", "C"), size = 14,
                  x = c(0, .49, 0), y = c(1,1, 0.33) )

#two.Q1.SC.plots <- {Q1.SC.short.mouse_scatter_boxplot + {Tot.SC_mouse.av_ALL} + plot_layout(ncol = 2, widths = c(1,1), heights = c(2,2)) }

@@

Make a figure for the diagram.  Import the straightened bivalents.. Load in the wsb double triangle.

@@

```{r Q1.M1_short_biv_lmer, echo=FALSE, warning=FALSE, message=FALSE}

lmer.Q1_SCLength_short_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Short.biv_mouse_table, REML = TRUE)

lmer.Q1_SCLength_short_M1_R.sex <- lmer(mean_SC ~ subsp + (1|strain), data=Short.biv_mouse_table, REML = TRUE)

lmer.Q1_SCLength_short_M1_R.subsp <- lmer(mean_SC ~ sex + (1|strain), data=Short.biv_mouse_table, REML = TRUE)

@@

We use the reduce data set of short bivalents to remove the effects of the paired X-X bivalent in females. The mouse average of short bivalents are significantly longer in females than males in all strains (t-test; p < 0.05) except _musculus^SKIVE^_, which my be attributed to low sample sizes (p = r Q1_mus.av_shortSC_t.test.SKIVE$p.value). The ratios of the mouse mean SC across sexes range across strains from `r Q1SC.mus.av.short.biv_DIFF_MSM` ( _musculus^MSM^_ ) to `r Q1SC.mus.av.short.biv_DIFF_WSB` ( _domesticus^WSB^_ ).

@@

Short.biv_mouse_table$subsp <- factor(Short.biv_mouse_table$subsp, ordered = FALSE)
Short.biv_mouse_table$sex <- factor(Short.biv_mouse_table$sex, ordered = FALSE)
Short.biv_mouse_table$strain <- factor(Short.biv_mouse_table$strain, ordered = FALSE)

Q1.glm_short.SC_M2 <- glm(mean_SC ~ subsp * strain * sex, data=Short.biv_mouse_table)

@@

@@

Q1.glm_short.SC_M3 <- glm(mean_SC ~ strain * sex, data=Short.biv_mouse_table)

@@

#make comparisons for the high quality cells
#get Q12 cells from MLH1.merge.Skel
#calq the mouse averages

@@

totSC_Q12_mouse.av.table <- ddply(.data= MLH1.merge.Skel %>% filter(quality < 3), 
                 .(mouse),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                # mean.MLH1 = mean( nMLH1.foci, na.rm = TRUE),
                # var.MLH1 = var( nMLH1.foci, na.rm = TRUE),

@@

#Long.biv_mouse_table

#mouse av
Q1.SC.long.mouse_scatter_boxplot <- ggplot( Long.biv_mouse_table, aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(75,150) )+ggtitle("Mouse average SC Length \n long bivalent dataset")

```{r Q1_SC.biv.long_ttest, echo=FALSE, include=TRUE}

@@

```{r Q1.F1pos_long.short_setup, echo=FALSE, eval=FALSE}
#move to setup

@@

```{r Comp_Ratios, echo=FALSE, eval=FALSE}
